---
title: "HW1_EC"
output:
  word_document: default
  html_document: default
---

################# Tournament selection

```{r}
tournament =  c ("0101110111","0111001001","0011001001","1110101100","1000010010","0000100010")
set.seed(42)

random_tournament = sample(tournament,6)
f = vector()

for (i in 1:length(random_tournament) ) {
      f[i]=  table(strsplit(random_tournament[i],"")) [2] 
}

mating_pool=  vector()

#slinding window of 3 elements

for (i in 1:length(f) ) {
  sliding_window =  seq(from=i,to=i+2,by=1)
  overflow=  i+2 -  length(f)
  if (overflow > 0) {
      sliding_window = c(sliding_window[c(1:(3 - overflow))],1:overflow)
  }
  mating_pool[i] = max(f[sliding_window] )
}


```
# • How many copies of each chromosome are present in the mating pool?

```{r}
table(mating_pool)

```


# • What is the average fitness of the chromosomes in the mating pool?
```{r}
mean(mating_pool)
```

# If the tournament size is reduced to one, what is the probability that the chromosome 1110101100
appears in the mating pool?

```{r}
#if the ournament size is reduced to one means there is no comptetition so it will be 1/n and in this case 1/6 probability to appear in the mating pool

```


# • If the tournament size is increased to five, what is the probability that the chromosome 0111001001
appears in the mating pool?


```{r}
tournament =  c ("0101110111","0111001001","0011001001","1110101100","1000010010","0000100010")
set.seed(42)

random_tournament = sample(tournament,6)
f = vector()

for (i in 1:length(random_tournament) ) {
      f[i]=  table(strsplit(random_tournament[i],"")) [2] 
}

mating_pool=  vector()

#slinding window of 3 elements

for (i in 1:length(f) ) {
  sliding_window =  seq(from=i,to=i+4,by=1)
  overflow=  i+4 -  length(f)
  if (overflow > 0) {
      sliding_window = c(sliding_window[c(1:(5 - overflow))],1:overflow)
  }
  mating_pool[i] = max(f[sliding_window] )
}

mating_pool

# 0 since it must the top 2 in oder to appear 
```


#################### Whole arithmetic crossover
```{r deffinition of the chromosomes}
x = c(0.23, 0.57, 0.29, 0.44, 0.44)

y = c(0.63, 0.82, 0.62, 0.15, 0.51)



```



```{r 0.5}
#ui= alpha* xi + (1-alpha)*yi
alpha= 0.5
u_0.5 = (alpha * x) + (1-alpha)*y
v_0.5= (alpha* y) + (1-alpha)*x
#results
u_0.5
v_0.5
```



```{r 0.1}
alpha= 0.1
u_0.1 = (alpha * x) + (1-alpha)*y
v_0.1= (alpha* y) + (1-alpha)*x
#results
u_0.1
v_0.1

```


################################### 3 Selection
```{r}
Population = c("AAOOEIIIOEA", "EEOUEO" ,"UUIOOAAIEEO","AAAEEEIIIOUUU","AUEIOUOOEEIIUIA")
fitness = vector() 
fitness_k = vector() 

for (i in 1:length(Population)){
  fitness[i] = sum(floor(table(strsplit(Population[i],""))/2))
  fitness_k[i] =  fitness[i] * 5
}

```

## • Calculate the probabilities of selecting each one of these individuals, based on a proportional selection
(based on the fitness).


```{r}
prop_selection= print(fitness / sum(fitness))
prop_selection= print(fitness_k / sum(fitness_k))

library(ggplot2)

selection_df <- data.frame(population = Population,prop_selection=prop_selection)
ggplot(selection_df, aes(x=population, y=prop_selection)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=round(prop_selection,5)), vjust=1.6, color="black", size=3.5)+
  theme_minimal()
```


## • Calculate the probabilities of selecting each one of these individuals, based on a linear ranking selection
with C = 2.

```{r}
r = vector()
r_k = vector()

c = 2
n_1 = (length(Population)-1)

for (i in 1:length(fitness) ){
  r[i] = which(sort(fitness) == fitness[i])[1] -1 
  r_k[i] = which(sort(fitness_k) == fitness_k[i])[1] -1 
}

linear_ranking_fitness = print(r * (c/n_1)) 
linear_ranking_fitness_k = print(r_k * (c/n_1)) 

selection_df <-cbind(selection_df, linear_ranking_fitness) 

ggplot(selection_df, aes(x=population, y=linear_ranking_fitness)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=round(linear_ranking_fitness,5)), vjust=1.6, color="black", size=3.5)+
  theme_minimal()

ggplot(selection_df, aes(x=population, y=linear_ranking_fitness_k)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=round(linear_ranking_fitness_k,5)), vjust=1.6, color="black", size=3.5)+
  theme_minimal()
```

## • Calculate the probabilities of selecting each one of these individuals, based on an exponential ranking
selection with m = 3.

```{r}
m = 3

exponential_ranking_fitness = print(m * (r/n_1) ^ (m-1)) 

exponential_ranking_fitness_k = print(m * (r_k/n_1) ^ (m-1)) 

ggplot(selection_df, aes(x=population, y=exponential_ranking_fitness)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=round(exponential_ranking_fitness,5)), vjust=1.6, color="black", size=3.5)+
  theme_minimal()

ggplot(selection_df, aes(x=population, y=exponential_ranking_fitness_k)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=round(exponential_ranking_fitness_k,5)), vjust=1.6, color="black", size=3.5)+
  theme_minimal()
```



```{r 1}
#ui= alpha* xi + (1-alpha)*yi
alpha= 1.0
u_1 = (alpha * x) + (1-alpha)*y
v_1= (alpha* y) + (1-alpha)*x
#results
u_1
v_1
```

```{r 0}
#ui= alpha* xi + (1-alpha)*yi
alpha= 0
u_0 = (alpha * x) + (1-alpha)*y
v_0= (alpha* y) + (1-alpha)*x
#results
u_0
v_0
```